### 2018-07-25 Thesis/Parts/Abstract/Abstract.tex
update

\documentclass[../../Thesis.tex]{subfiles}
\begin{document}
Over the last few years, word embeddings have taken a dominant position in the Information Retrieval domain. Many studies have been done concerning the quality and application of word embeddings on general texts, such as the Wikipedia corpus and comments on review websites. Giving promising results, the word embeddings have been studied and improved over recent years. However, these studies have been focused on generic texts, which are not limited to the characteristics of in-domain texts such as rare domain-specific words or have been focussed on small sets of academic texts. This research focusses on the quality and application of word embeddings on domain-specific texts, concerning a large corpus of 1.391.543 scientific articles which have been published in 2017. We validate the word embeddings using a categorization task to match articles to journals. We furthermore create a 2-dimensional visualization of the word embeddings on journal level, to visualize journal relatedness.\\
\vspace*{\fill}
\\
We aim to have this thesis published in the proceedings of the \textit{ECIR} (\url{http://ecir2019.org/}). \textit{Submission deadline: 9-10-2018}.
\end{document}

>>>>

\documentclass[../../Thesis.tex]{subfiles}
\begin{document}
Over the last few years, word embeddings have taken a dominant position in the information retrieval research. Studies conducted have focused on the quality and application of word embeddings on corpuses such as Wikipedia or online comments/reviews/tweets. However, these studies are limited to generic texts and lack technical, scientific or domain specific nuances such as rare words, lots of abbreviations, or chemical/mathematical formulas commonly used in academic context. This research focusses on the quality and application of word embeddings on academic corpuses. Word embedding requires less memory and is processed faster than conventional alternatives such as TFIDF. Hence, we intend to use word embeddings as an efficient way to model the content for our recommendation/disambiguation/search engines. This study focuses on word embedding compared to TFIDF for modeling content in scientific articles. We use a word2vec model  trained on titles and abstracts of roughly 68 millions scientific articles. Word embedding are evaluated against existing benchmarks for generic texts from previous studies. Furthermore, we have developed a new benchmark to evaluate these embedding in the scientific context. We validate the word embeddings via a categorization task to match articles to journals for roughly 1.4 millions articles published in 2017. Our results show that word embedding is a better content model for titles (short text) while TFIDF is a better model for abstracts (longer text). Furthermore, we have created a 2-dimensional visualization of the journals modeled via embeddings to visualize their relatedness. This graphs can be used to find competitive journals or gaps to propose new journals.\\
\vspace*{\fill}
\\
We aim to have this thesis published in the proceedings of the \textit{ECIR} (\url{http://ecir2019.org/}). \textit{Submission deadline: 9-10-2018}.
\end{document}

### 2018-07-25 Thesis/Parts/Background/Background.tex
update

\documentclass[../../Thesis.tex]{subfiles}
\begin{document}
\header{Introduction and Background}
\subheader{Information retrieval}
Information Retrieval(IR)is the activity of gathering relevant information, given another initial piece of information. 
\\\\The most practical example of this is a search engine. Given one or more search words (a query) the search engine will attempt to find relevant information. For example, an online search for "Information Retrieval" (initial piece of information) will give you a list of results (relevant information). To be able to do this, the search engine must know which texts are related. \\\\Interpreting which texts are related can be achieved with, and without neural networks, a traditional technique which does not use a neural network is TF-IDF, shot for Term Frequency - Inverted Document Frequency. Techniques that use a neural network are (among others) Word2Vec, Paragraph Vectors and GloVe.

\subheader{Neural Network}
A complete in-depth background into neural networks is beyond the scope of this thesis, therefore we will only describe the simplified working of a neural network and its basic application in IR.
Neural Networks are multi-layered computational software, based on matrix transformations, which tries to map input values to output values. This mapping uses a pre-defined amount of layers that modify the values through matrix transformations. Neural networks rely on training to create optimal values for these layers. Finding these optimal values is done iteratively, via a process known as "back-propagation". Back-propagation is done by starting at the output layer of the Neural Network and tracking error in the output back through the layers. This error information is used to improve the values of the layers. The values in the layers are only changed at training time; high-quality (training) data is, therefore, essential for neural networks. Neural networks can be applied to many different tasks, as long as there is sufficient data available to create a training and validation set. The training is the process of iteratively adjusting multiplication matrices or vectors values to achieve the optimal result. Which is the result that is as close as possible to the given output for all input and output sets, without underperforming on other sets\footnote{Having only good results for the training/validation set is known as overfitting. This means that the layer-values are fine-tuned to only perform (extremely) well on one set while underperforming on others}.\\Neural Networks are used, in the creation of word embeddings, to either predict words that may occur around a given word or predict a word given words that surround it. For example, given the sentence\\
\begin{center}
\textit{"The quick brown fox jumps over the lazy dog"}
\end{center}
the Neural Network can be trained to either predict the words around "jumps", which we refer to as context words (the, quick, brown, fox, over, the, lazy, dog). It can also predict the word "jumps", given the context words (the, quick, brown, fox, over, the, lazy, dog). This results in a matrix, a vector per word which is referred to as a word embedding. This embedding indicates "word relatedness" which, as mentioned earlier enables association (thus retrieval) with related(relevant) texts. It is to be noted that the word embeddings are only able to represent word relatedness to the other words trained in the same run. Due to the random initialization of the initial values of the layers in the neural networks, they do not produce the same result every run. Therefore, word embeddings from different runs of the same neural network cannot be compared.

\subheader{Embedding}
An embedding is a distributed, numerical representation of text in a multi-dimensional vector space\footnote{The vector spaces of separately trained word embeddings differ, since each run the initial values of the neural network are randomly initialized. This means that the same word trained in two separate runs do not have to have the same embedding. However, they will have the same relationship to other words trained in their respective runs.} which can capture both the semantic and syntactic information\cite{mikolov2013distributed}. In the case of a word embeddings, the embeddings represent the words. These embeddings are created using machine learning models, which do not need human interaction~\cite{lai2016generate}, they are so-called \textit{unsupervised learning algorithms}. Once trained, the embeddings can also be used to construct embeddings for collections of texts. For example, a word embedding can be used to create a sentence, paragraph, document or corpus embeddings. The usage of word embeddings has improved various Natural Language Processing areas such as named entity recognition, part-of-speech tagging, parsing, and semantic role labelling~\citet{luong2013better}.

\subheader{Text analysis techniques}
To enable a computer to process text, for embedding creation or other tasks, the text has to be processed by an algorithm. In this research, we used embeddings created by Word2Vec and TF-IDF feature vectors.
\begin{jumpin}
\subsubheader{Word2Vec}
Word2vec word embeddings are created using a neural network, Word2Vec learns word embeddings via maximizing the log conditional probability of the word given the context word(s) occurring within a fixed-sized window. Therefore the learned embeddings contain useful knowledge about word co-occurrence\cite{nalisnick2016improving}. There are multiple input/output possibilities for the neural network, best known are Skip-gram and the Continuous Bag-of-Words model (CBOW). The Skip-gram model takes a target word as input and outputs the predicted context words, while CBOW takes the context words as input and outputs the predicted target word\cite{nalisnick2016improving, pennington2014glove}. This is illustrated in Figure~\ref{figure:word2vecModels}, in this Figure, $v(w)$ represents the target word, and $v(w...)$ represent the context words. \citet{mikolov2013distributed}\cite{mikolov2013efficient} presented several extensions to the word2vec model that improve the quality of vector training and its speed, such as introducing Hierarchical Softmax, Negative Sampling and the subsampling of frequent words to the Skip-gram approach. Variations to the word2vec model have also been proposed, such as the doc2vec model described by \citet{lau2016empirical} which creates document embeddings instead of word embeddings.
\begin{center}
\begin{figure}[hbt]
\includegraphics[width=4in]{Plots/Model-Architecture-of-CBOW-and-Skip-Gram.png}
\caption{Illustration of the CBOW and Skip-Gram models, as presented by \citet{chen2017convolutional}. Left part of the Figure illustrates the CBOW model, context words are represented at the left side (input) and the center word (output) is represented as $v(w)$ as output. The right side of the Figure illustrates the Skip-Gram Model, which takes the center word as input (left) and outputs the context words (right).}\label{figure:word2vecModels}
\end{figure}
\FloatBarrier
\end{center}
\subsubheader{Paragraph vectors}
Variations on the word2vec model have also been proposed,~\citet{le2014distributed} introduced the paragraph vector in their paper "Distributed representations of sentences and documents". The Paragraph Vector framework is based on the word vectors framework. The difference between the frameworks is the calculation of the probability, the Paragraph Vector framework uses a matrix, which consists of every paragraph. This matrix is used to replace a concatenation or average of word vectors. An advantage of the paragraph vector model is that it takes the word order into consideration, at least in a small context \cite{le2014distributed}. \citet{dai2015document} state that the Paragraph Vectors model performs significantly better than other models on grouping, triplet finding and related object/article finding tasks on Wikipedia and arXiv texts.

\subsubheader{GloVe}
\citet{pennington2014glove} introduced the Global Vectors (GloVe) model. This model captures the global corpus statistics. The model transforms the word co-occurrences of all words in the corpus to chances, it excludes all the zero values and uses that as initial input for the neural network. This model out-performs other models on word analogy, word similarity and entity recognition according to the findings by~\citet{pennington2014glove}.\\

>>>>

\documentclass[../../Thesis.tex]{subfiles}
\begin{document}
\header{Introduction and Background}
\subheader{Academic articles}
Academic articles are text, focussed on specific domains (i.e. biology, economy or computer science), that contain scientific or domain specific nuances such as: rare words, abbreviations, or mathematical formulas. Because these text contain information specific to a domain, we call the articles "domain-specific"; meaning that they give information about a topic in a domain (i.e. biology). Articles are published in journals, which are collections of articles which share a common topic, i.e. a specific field inside the biology domain. Our research focusses on the academic articles and journals that are not limited to a specific domain.
\subheader{Information retrieval}
Information Retrieval (IR) is the activity of gathering relevant information, starting from an initial piece of information.  The most practical example of IR is a search engine. Given one or more search words (a query) the search engine will attempt to find relevant information. For example, a search for "Information Retrieval" (initial piece of information) will return a list of results (relevant information). To be able to return a list of results, the search engine must know which texts are related. \\Interpreting which texts are related can be achieved with, and without neural networks, a traditional technique which does not use a neural network is TF-IDF, short for Term Frequency - Inverted Document Frequency. Techniques that use a neural network are Word2Vec, Paragraph Vectors and GloVe.
\subheader{Neural Networks}
A complete in-depth background into neural networks is beyond the scope of this thesis, therefore we will only describe the simplified working of neural networks and their basic application in IR.
Neural Networks are mathematical models\cite{funahashi1989approximate}, which use matrix transformations, to map input values to output values. This mapping uses a pre-defined amount of layers that modify the values through matrix transformations. Neural networks rely on training to create optimal values for these layers. Finding these optimal values is done iteratively, via a process known as "back-propagation". Back-propagation is done by starting at the output layer of the neural network and tracking the error in the output back through the layers. This error information is used to improve the values of the layers. The values of the layers are only changed at training time; high-quality (training) data is, therefore, essential for neural networks\cite{Truong2017Thesis, lai2016generate}. Neural networks can be applied to many different tasks, as long as there is sufficient data available to create training and validation sets. During the training process, one seeks to achieve the optimal results, i.e. the result that is as close as possible to the given output for all input and output sets, without underperforming on other sets\footnote{Having only good results for the training/validation set is known as overfitting. This means that the layer-values are fine-tuned to only perform (extremely) well on one set while underperforming on others.}.\\Neural Networks are used, in the creation of word embeddings, to either predict words that may occur around a given word or predict a word given words that surround it. For example, given the sentence:\\
\textit{"The quick brown fox jumps over the lazy dog"}
a neural network can be trained to either predict the words around "jumps", which we refer to as context words (in this example, the context words are: the, quick, brown, fox, over, the, lazy, dog), or to predict the word "jumps", given the context words. The training process results in a matrix, which is transformed to a collection of vectors, each vector is associated to a word, these vectors are referred to as word embeddings. This embedding indicates "word relatedness" with other word embeddings (representing other words) which, as mentioned earlier, enables association (thus retrieval) with related (relevant) texts. It is to be noted that the word embeddings are only able to represent word relatedness to the other words trained in the same session. Due to the random initialization of the initial values of the layers in neural networks, the neural networks do not produce the same result every run. Therefore, word embeddings from different runs of the same neural network cannot be compared.
\subheader{Embedding}
An embedding is a distributed, numerical representation of text in a multi-dimensional vector space\footnote{The vector spaces of separately trained word embeddings differ, since each run the initial values of the neural network are randomly initialized. This means that the same word trained in two separate sessions may have two different embeddings. However, they will have the same relationship to other words trained in their respective sessions.} which can capture both semantic and syntactic information\cite{mikolov2013distributed}. In the case of word embeddings, the embeddings represent the words. These embeddings can be\footnote{This depends on the method used to create the word embeddings, for this research, we used word embeddings trained by an unsupervised learing algorithm.} created using machine learning models, which do not need human interaction~\cite{lai2016generate}, they are so-called \textit{unsupervised learning algorithms}. Once trained, the embeddings can also be used to construct embeddings for collections of texts. For example, a word embedding can be used to create sentence, paragraph, document or corpus embeddings. The usage of word embeddings has improved various natural language processing (NLP) areas such as named entity recognition, part-of-speech tagging, parsing, and semantic role labelling~\cite{luong2013better}. Embeddings can furthermore be used in search and recommendation. Potential application of the embeddings in the academic domain could be the improvement of search engines, text analysis to improve NLP tasks for academic texts, and journal recommendation for publishing articles. The embeddings can furthermore be used to help researchers which are new to a field to explore the field by finding related articles or articles that are related to the field of expertise of the researcher and the new research field. Embeddings can also be visualized (see paragraph 1.6), this can give insight into the topics of popular journals, and can be used to indicate research subjects where there are few publications or journals.
\subheader{Text analysis techniques}
To enable a computer to process text, for embedding creation or other tasks, the text has to be processed by an algorithm. In this research, we used embeddings created by Word2Vec, and feature vectors\footnote{Which do not require training (see subparagraph TF-IDF).} created using TF-IDF.
\subsubheader{Word2Vec}
Word2vec word embeddings are created using a neural network, Word2Vec learns word embeddings via maximizing the probability of the word given the context word(s) occurring within a fixed-sized window, therefore the learned embeddings contain useful knowledge about word co-occurrence\cite{nalisnick2016improving} which indicates relatedness\cite{lai2016generate}. There are multiple input/output possibilities for the neural network. The best known are Skip-gram and the Continuous Bag-of-Words model (CBOW). The Skip-gram model takes a target word as inp ut and outputs the predicted context words, while CBOW takes the context words as input and outputs the predicted target word\cite{nalisnick2016improving, pennington2014glove}. This is illustrated in Figure~\ref{figure:word2vecModels}, where, $v(w)$ represents the target word, and $v(w...)$ represent the context words. \citet{mikolov2013distributed}\cite{mikolov2013efficient} presented several extensions to the word2vec model that improve the quality of the embeddings and the training speed, such as Hierarchical Softmax, Negative Sampling and the subsampling of frequent words for the Skip-gram approach. Variations to the word2vec model have also been proposed, such as the doc2vec model described by \citet{lau2016empirical} which creates document embeddings\footnote{Document embeddings are embeddings which capture the relatedness between documents.} instead of word embeddings.
\includegraphics[width=4in]{Plots/Model-Architecture-of-CBOW-and-Skip-Gram.png}
\caption{Illustration of the CBOW and Skip-Gram models, as presented by \citet{chen2017convolutional}. Left part of the Figure illustrates the CBOW model, context words are represented at the left side (input) and the center word (output) is represented as $v(w)$ as output. The right side of the Figure illustrates the Skip-Gram Model, which takes the center word as input (left) and outputs the context words (right).}\label{figure:word2vecModels}
\includegraphics[width=4in]{Plots/training_window.png}
\caption{Illustration of the training windows. Illustration created by \citet{trainingWindow}.}
\subsubheader{Paragraph vectors}
Variations on the word2vec model have also been proposed,~\citet{le2014distributed} introduced the paragraph vector in their paper "Distributed representations of sentences and documents". The Paragraph Vector framework is based on the word vectors framework. The difference between the frameworks is the calculation of the probability, the Paragraph Vector framework uses a matrix, which consists of every paragraph. This matrix is used to replace a concatenation or average of word vectors. An advantage of the paragraph vector model is that it takes the word order into consideration, at least in a small context \cite{le2014distributed}. \citet{dai2015document} state that the Paragraph Vectors model performs significantly better than other models on grouping, triplet finding and related object/article finding tasks on Wikipedia and arXiv texts.

\subsubheader{GloVe}
\citet{pennington2014glove} introduced the Global Vectors (GloVe) model. This model captures the global corpus statistics. The model transforms the word co-occurrences of all words in the corpus to probabilities, it excludes all the zero probabilities and uses that as initial input for the neural network. This model outperforms other models on word analogy, word similarity and entity recognition according to the findings by~\citet{pennington2014glove}.\\
